---
title: "AccuracyOFmovement"
author: "CF Hainey"
date: "August 24, 2016"
output: 
  html_document:
    keep_md: true
---
#Introduction
Technology and sports are merging together in complex and insightful ways.  You're every movement, pulse, calorie burn, stress level, quality and quantity of your sleep can be monitored with a watch.  The Apple watch is also totted to monitor your diabetes.  For all of these aspects, you are given quantity.  Only sleep is there quality data recorded.  I have written a script to analyze the quality of movement on 6 individuals.  These individuals were asked to lift dumbbells in 5 different ways, from proper to sloppy.  My task was to forecast how 20 different measurements were performed.

I built my model by finding the least amount of variables to explain my dataset using PCA analysis with the package FactoMineR and FACToextra. I found user name, num window, roll belt, pitch belt,total accel belt, and yaw belt to describe the dataset the best.  I performed this PCA analysis. 

Next, I used only these six variables identified in the PCA anlysis to run randomForester on a subset of the training data.  With a high accuracy, I applied the randomForest model to the final test set given by the instructor.  With the test set, the out of sample error was around 1%. I choose this method for it high accuracy and ease of use with RandomForest and caret packages.


#Background
Data came from http://groupware.les.inf.puc-rio.br/har., and the instruction for goal came from the Coursera's Practical Machine Learning course websit https://www.coursera.org/learn/practical-machine-learning/supplement/PvInj/course-project-instructions-read-first.  More Background informatino can be found at http://groupware.les.inf.puc-rio.br/har.

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=FALSE,message=FALSE}

#load data
library(data.table) # allows you to brind in specific columns
dat <- fread("./training1.csv", select = c('user_name','num_window','roll_belt','pitch_belt','yaw_belt','total_accel_belt', 'gyros_belt_x', 'gyros_belt_y', 'gyros_belt_z','accel_belt_x','accel_belt_y','accel_belt_z', 'magnet_belt_x', 'magnet_belt_y', 'magnet_belt_z', 'roll_arm', 'pitch_arm', 'yaw_arm','total_accel_arm', 'gyros_arm_x', 'gyros_arm_y','gyros_arm_z', 'accel_arm_x','accel_arm_y', 'accel_arm_z','magnet_arm_x', 'magnet_arm_y', 'magnet_arm_z', 'roll_dumbbell', 'pitch_dumbbell', 'yaw_dumbbell','total_accel_dumbbell', 'gyros_dumbbell_x',  'gyros_dumbbell_y',  'gyros_dumbbell_z', 'accel_dumbbell_x','accel_dumbbell_y', 'accel_dumbbell_z', 'magnet_dumbbell_x', 'magnet_dumbbell_y', 'magnet_dumbbell_z', 'roll_forearm', 'pitch_forearm', 'yaw_forearm', 'total_accel_forearm', 'gyros_forearm_x', 'gyros_forearm_y', 'gyros_forearm_z', 'accel_forearm_x',  'accel_forearm_y', 'accel_forearm_z','magnet_forearm_x', 'magnet_forearm_y','magnet_forearm_z','classe'))

#change names to dummy variables
dat$user_name[dat$user_name == 'carlitos'] <- 1
dat$user_name[dat$user_name == 'pedro'] <- 2
dat$user_name[dat$user_name == 'adelmo'] <- 3
dat$user_name[dat$user_name == 'charles'] <- 4
dat$user_name[dat$user_name == 'eurico'] <- 5
dat$user_name[dat$user_name == 'jeremy'] <- 6
dat$user_name <-as.numeric(dat$user_name)

set.seed(1234)
#subset data for validation
library(ggplot2); library(caret);
# Create a building data set and validation set
inBuild <- createDataPartition(y=dat$classe, p=0.7, list=FALSE)
validation <- dat[-inBuild,]; buildData <- dat[inBuild,]
inTrain <- createDataPartition(y=buildData$classe, p=0.7, list=FALSE)
training <- buildData[inTrain,]; testing <- buildData[-inTrain,]
train1 <- training

# set up PCA
library("factoextra")
library("FactoMineR")
train1 <- training
pca_set<-train1[,classe:=NULL]
res.pca <- PCA(pca_set, graph = FALSE)
#print(res.pca)
eigenvalues <- res.pca$eig
head(eigenvalues[, 1:2])
fviz_screeplot(res.pca, ncp=10)
# Coordinates of variables
head(res.pca$var$coord)
fviz_pca_var(res.pca)

#load only values found in pca analysis in.
library(data.table) # allows you to brind in specific columns
dat <- fread("./training1.csv", select = c( "user_name", "num_window", "roll_belt", "pitch_belt", "yaw_belt", "total_accel_belt", "classe" ))
test2 <- fread("./testing1.csv", select = c( "user_name", "num_window", "roll_belt", "pitch_belt", "yaw_belt", "total_accel_belt", "classe"))

#change names to dummy variables and give columns proper class
dat$user_name[dat$user_name == 'carlitos'] <- 1
dat$user_name[dat$user_name == 'pedro'] <- 2
dat$user_name[dat$user_name == 'adelmo'] <- 3
dat$user_name[dat$user_name == 'charles'] <- 4
dat$user_name[dat$user_name == 'eurico'] <- 5
dat$user_name[dat$user_name == 'jeremy'] <- 6
test2$user_name[test2$user_name == 'carlitos'] <- 1
test2$user_name[test2$user_name == 'pedro'] <- 2
test2$user_name[test2$user_name == 'adelmo'] <- 3
test2$user_name[test2$user_name == 'charles'] <- 4
test2$user_name[test2$user_name == 'eurico'] <- 5
test2$user_name[test2$user_name == 'jeremy'] <- 6
test2$num_window <-as.numeric(test2$num_window)
test2$total_accel_belt <-as.numeric(test2$total_accel_belt)
test2$user_name <-as.factor(test2$user_name)
set.seed(1234)
#run modeling and prediction
library(ggplot2); library(caret);
# Create a building data set and validation set
inBuild <- createDataPartition(y=dat$classe,p=0.7, list=FALSE)
validation <- dat[-inBuild,]; buildData <- dat[inBuild,]
inTrain <- createDataPartition(y=buildData$classe,p=0.7, list=FALSE)
training <- buildData[inTrain,]; testing <- buildData[-inTrain,]
train1 <- training
training$num_window <-as.numeric(training$num_window)
training$total_accel_belt <-as.numeric(training$total_accel_belt)
training$user_name <-as.factor(training$user_name)
testing$num_window <-as.numeric(testing$num_window)
testing$total_accel_belt <-as.numeric(testing$total_accel_belt)
testing$user_name <-as.factor(testing$user_name)
set.seed(1234)
library(randomForest)
modrf = randomForest(factor(classe) ~ ., data=training, na.action=na.roughfix,importance=TRUE)
modrf

#predict with test and testing2 set
importance(modrf, type = 2)
forest.pred <- predict(modrf, testing)
forest.per <- table(testing$classe, forest.pred, dnn=c("Actual", "Predicted"))
forest.per
forest_val <-predict(modrf, test2)
forest_val
```

#Resources
http://www.sthda.com/english/wiki/factominer-and-factoextra-principal-component-analysis-visualization-r-software-and-data-mining
